### 2016-06-15 inst/doc/Genus.Rmd

---
title: "Genus"
author: "Sylvain Schmitt"
date: "May 26, 2016"
output: html_document
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 10, fig.width = 10,
  cache = T, cache.lazy = F)
```

# Data
```{r Data, echo=TRUE}
rm(list = ls()); invisible(gc())
library(Uppangala)
library(raster)
Env <- genEnv()
Trees <- genTrees()
Species <- genSpecies()
Trees$SpCode <- as.factor(tolower(as.character(Trees$SpCode))) # Should be treated in gen functions
row.names(Species) <- tolower(row.names(Species))
# Inconsistancies
sort(Species$SpCode)[which(!(sort(Species$SpCode) %in% sort(levels(Trees$SpCode))))]
sort(levels(Trees$SpCode))[which(!(sort(levels(Trees$SpCode)) %in% (sort(Species$SpCode))))]
```

# Species plots
```{r Plot sp, echo=FALSE}
SpCode <- sort(Species$SpCode)[which(sort(Species$SpCode) %in% sort(levels(Trees$SpCode)))] # Using only common SpCode
Genus <- unlist(lapply(strsplit(Species[SpCode,2], ' ', fixed = T), function(x){x[[1]]}))
for(Gen in unique(Genus)){
  if(length(which(Genus == Gen)) > 1){
    main <- paste(Gen, Species[which(Genus == Gen)[1],3], sep = ', ')
    plot(Env$DEM, main = main, col = grey(100:-1:0/100), alpha = 0.8)
    contour(Env$DEM, add = T)
    col <- rainbow(length(which(Genus == Gen)))
    j <-  1
    Sp.l <- c()
    for(i in which(Genus == Gen)){
      Sp <-  SpCode[i]
      Sp.l <- c(Sp.l, Sp)
      points(Y ~ X, Trees[Trees$SpCode == Sp,], pch = 20, col = col[j])
      j <-  j + 1
    }
    species.v <- unlist(lapply(strsplit(Species[Sp.l,2], ' ', fixed = T), function(x){x[[2]]}))
    legend('bottomright', species.v, fill = col)
  }
}
```