### 2018-07-09 draft.Rmd
Uppangala 0.1.9047 Bulletproofing code & ToDo.md

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt, Maxime Réjou-Méchain, Valérie Raevel, Ayyappan N, Balachandran N., Barathan N., and François Munoz"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
linestretch: 1.5
csl: bibliography/mee.csl
bibliography: bibliography/refs.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(parallel)
library(raster)
library(broom)
library(vegan)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggfortify)
sapply(list.files("R", full.names = T), source)
theme_set(bayesplot::theme_default())
cores <- detectCores() - 1
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
```

# Traits

```{r Traits}
genPFT() %>% 
  mutate(id = row.names(.)) %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(id.vars = "id", variable.name = 'Abbreviation') %>% 
  mutate(Abbreviation = as.character(Abbreviation)) %>% 
  left_join(data.frame(
    Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
    Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
    Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
    Strategy = c('Leaf defense and investment',
                 'Leaf resource capture vs.investment',
                 'Leaf defense and investment',
                 'Leaf resource acquisiton vs. defense',
                 'Stem transport, structure and defense'),
    stringsAsFactors = F
  ),  by = "Abbreviation") %>% 
  group_by(Trait, Abbreviation, Unit, Strategy) %>% 
  summarise_if(is.numeric, funs(min(.,na.rm = T), max(., na.rm = T))) %>% 
  mutate(range = paste(round(min,2), "-", round(max,2))) %>% 
  select(-min, -max) %>% 
  left_join(genPFT() %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'Abbreviation', 
                   value.name = 'Intra-individuals') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Tree, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'Abbreviation', value.name = 'Intra-species') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Sp, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>%
              mutate(id = row.names(.)) %>% 
              melt(id.vars = "id", 
                   variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
              select(-id) %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Abbreviation) %>%
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))),
            by = "Abbreviation") %>%
  mutate(`Intra-individuals` = ifelse(Abbreviation == "WD", NA, `Intra-individuals`)) %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc',
        caption = 'Functional traits measured. Traits with their abbreviation, standard unit, associated species strategy, range of values, intra-individual, intra-specific and inter-specific component of variation.  Species strategy are precised for functioning, survival and reproduction [@baraloto_decoupled_2010]. The rlast columns represents the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species.')
```

```{r Q}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  group_by(Tree, Sp, SpCode, CE) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree, -CE) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Sp, -LTD, -SHA)
suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
Q <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(data.frame(PFT.pca$scores) %>% 
              mutate(SpCode = row.names(.)),
            by = "SpCode") %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata))
```

```{r L}
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
L <- genTrees() %>% 
  mutate(com = (XY %over% com)[,1]) %>% 
  filter(!is.na(com)) %>% 
  group_by(com, SpCode) %>% 
  summarise(n = n()) %>% 
  dcast(com ~ SpCode, value.var = "n") %>% 
  replace(is.na(.), 0)
```

```{r R}
# Env <- genEnv()
# com <- quadrats(cs = c(30,30))
# com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'), path = '/usr/share/applications')
# names(Deriv)[5] <- 'SW'
# com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
# com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
# BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
# com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
# pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
# save(pca.env, file = file.path(path, 'pca.env.Rdata'))
# R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
# row.names(R) <- com@data$id
# R$community <- row.names(R)
# save(R, file = file.path(path, 'R.Rdata'))
load(file.path("draft_save", 'R.Rdata'))
```

# PCA

```{r PCA,  fig.cap='Principal Component Analysis (PCA) of species mean trait values. Marginal plots represent the density distribution of species on the two principal component axis for each morphotype. See table \\@ref(tab:Traits) for abbreviation.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'lightgrey') +
  geom_vline(aes(xintercept = 0), col = 'lightgrey') +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1))
spectrum.density <-pca.plot$data %>% 
  select(Comp.1, Comp.2, strata) %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  melt(id.vars = "strata", variable.name = "spectrum") %>% 
  ggplot(aes(value, color = spectrum, fill = spectrum)) +
  geom_density(alpha = 0.5) + 
  facet_wrap(~strata, nrow = 2) +
  coord_flip()
cowplot::plot_grid(pca.plot, spectrum.density, rel_widths = c(2,1))
```

```{r varianceTest}
data <- pca.plot$data %>% 
  select(Comp.1, Comp.2, strata) %>% 
  # rename(WES = Comp.1, LES = Comp.2) %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% # ! test !
  melt(id.vars = "strata", variable.name = "spectrum")
Variance <- list(
  LES = list(
    C = round(var(data[data$strata == "C" & data$spectrum == "LES", "value"], na.rm = T), 3),
    U = round(var(data[data$strata == "U" & data$spectrum == "LES", "value"], na.rm = T), 3)),
 WES = list(
   C = round(var(data[data$strata == "C" & data$spectrum == "WES", "value"], na.rm = T), 3),
   U = round(var(data[data$strata == "U" & data$spectrum == "WES", "value"], na.rm = T) ,3))
)
Leven_test <- list(
  C = round(car::leveneTest(value ~ spectrum, data[data$strata == "C",])$`Pr(>F)`[1], 3),
  U = round(car::leveneTest(value ~ spectrum, data[data$strata == "U",])$`Pr(>F)`[1], 3)
)
```

Canopy species vary significantly more on the wood than the leaf economic spectrum (`r Variance$LES$C` vs `r Variance$WES$C`, Levene's test p-value=`r Leven_test$C`). In parrallel, understorey species show a tendency to vary more on the leaf than the wood economic spectrum (`r Variance$LES$U` vs `r Variance$WES$U`, Levene's test p-value=`r Leven_test$U`).

# Guilds

```{r GWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(..p.signif..))) 
```

# LM

```{r GWMses}
GWM <- L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(Q, by = "SpCode") %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
#   return(x)
#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     melt(id.var = "com",
#          variable.name = "SpCode",
#          value.name = "abundance") %>%
#     dcast(SpCode ~ com) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
load(file.path("draft_save", 'nGWM.Rdata'))
GWM_ses <- GWM
GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
```

```{r LM}
GWM_ses %>% 
  filter(!is.na(strata)) %>% 
  melt(id.vars = c('strata', 'community')) %>% 
  rename(trait = variable) %>% 
  rename(SES = value) %>% 
  mutate(community = as.character(community)) %>% 
  left_join(R, by = "community") %>%
  group_by(strata, trait) %>% 
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                     direction = 'both', trace = 0)) %>% 
  tidy(reg) %>% 
  group_by(strata, trait, term) %>%
  transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
                          stars(p.value, ns = '   '))) %>% 
  dcast(strata + trait ~ term) %>% 
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
  mutate_all(funs(replace(., is.na(.), ' '))) %>% 
  left_join( # Inclusion of R2 but should be split
    GWM_ses %>% 
      filter(!is.na(strata)) %>% 
      melt(id.vars = c('strata', 'community')) %>% 
      rename(trait = variable) %>% 
      rename(SES = value) %>% 
      mutate(communtiy = as.character(community)) %>% 
      left_join(R, by = "community") %>%
      group_by(strata, trait) %>% 
      do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                    direction = 'both', trace = 0)) %>% 
      glance(reg) %>% 
      select(r.squared) %>% 
      mutate(r.squared = round(r.squared, 2)) %>% 
      rename("$R^2$" = r.squared), by = c("strata", "trait")
  )%>% # End of R2
  filter(trait != 'LES') %>% 
  filter(trait != 'WES') %>%
  bold() %>% 
  kable(caption = 'Linear models of the standardized effect size (SES) with null model NM2 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. We performed stepwise selection of the predictors including slope, curvature, wetness southwesterness (SW), and canopy height (Canopy), and only the selected predictors are shown. See table \\@ref(tab:Traits) for abbreviations. Bold and number of asterisk indicates the significance threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001).', format = 'pandoc')
```

# Environment

```{r PCAenv, fig.cap='Principal component analysis (PCA) performed on topgraphical variables. SW stand for southwesterness.'}
load(file.path("draft_save", 'pca.env.Rdata'))
pca.env %>% 
  autoplot(colour = 'grey', loadings.label.size = 6,
           loadings.label.colour = 'black',
           loadings = T, loadings.label = T, 
           loadings.colour = 'black',
           loadings.label.vjust = 2) +
  coord_equal() +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  xlab('Axe 1 − 42.52 %') + ylab('Axe 2 − 23.44 %') 
```

# Why not hypervolumes

```{r hypervolumes}
# library(hypervolume)
# bd <- Q %>% 
#   filter(!is.na(WES)) %>% 
#   select(-LES, -WES, -SpCode, -strata) %>% 
#   summarise_all(funs(sd)) %>% 
#   unlist()
# hv <- hypervolume_set(hypervolume(na.omit(Q[Q$strata == "C",names(bd)]), 
#                                   kde.bandwidth = bd, verbose = F), 
#                       hypervolume(na.omit(Q[Q$strata == "U",names(bd)]), 
#                                   kde.bandwidth = bd, verbose = F), 
#                       check.memory = F, verbose = F)
# plot(hv)
# rm(hv)
```

* Outliers make quicly overlapping hypervolumes besides the density stay low in those area
* Hypervolumes consider uncorrelated axis, which is not true for functional traits, and blur strategy axes highlighted by economic spectra