### 2020-01-23 Manuscript.Rmd
Uppangala 0.1.9057 23.20 MRM & VR review

---
title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Sp, -LTD, -SHA)
suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
Q <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(data.frame(PFT.pca$scores) %>% 
              mutate(SpCode = row.names(.)),
            by = "SpCode") %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata))
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
L <- genTrees() %>% 
  mutate(com = (XY %over% com)[,1]) %>% 
  filter(!is.na(com)) %>% 
  group_by(com, SpCode) %>% 
  summarise(n = n()) %>% 
  dcast(com ~ SpCode, value.var = "n") %>% 
  replace(is.na(.), 0)
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWMses
GWM <- L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(Q, by = "SpCode") %>% 
  filter(!is.na(strata)) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
#   return(x)
#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(detectCores())
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     melt(id.var = "com",
#          variable.name = "SpCode",
#          value.name = "abundance") %>%
#     dcast(SpCode ~ com) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
load(file.path("draft_save", 'nGWM.Rdata'))
GWM_ses <- GWM
GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
```

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - email
    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

\newpage

# Abstract

1. Topographic heterogeneity determines changing soil conditions, and environment also changes across vertical strata in rainforests. However, the joint influence of these factors on tree functional composition remain little known.
<!-- Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.  -->
<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.  -->
<!-- Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests. -->
We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.  
1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. 
Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

## Keywords

<!-- > To be checked at the end -->
Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

\newpage

# Introduction

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Specifically, topography is known to filter different ecological strategies [@kraft_functional_2008] due to its impact on water and nutrient availability [@ferry2010higher; @Blanchard2019].
<!-- At higher topographic positions, e.g., hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review]. -->
However, tree species can occupy different vertical strata at adult stage, and thus be classified into vertical guilds (understory and canopy), experimenting contrasted environmental conditions. 
Indeed, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical guilds is little addressed in tropical forests [@wright2002plant].

<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. -->

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes. Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. 
Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
However, functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. 
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].

<!-- Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds. Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].  -->
In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review] of canopy trees to water and nutrient limitation on hilltops [@ferry2010higher], 
 <!-- In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), -->
while limited light availability should favor conservative strategies in understory trees (along the leaf economics spectrum). Thus we expected different trait-environment relationships across vertical guilds, depending on tree species strategies along the fundamental leaf and wood economics spectra. Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].

We assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations. We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000]. The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play an important role in tree community assembly.

We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? (ii) Does functional composition differ between assemblages of canopy and understory guilds? and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?

<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species. We hypothesized functional composition to differ between canopy and understory species, due to their differences in growth and functional strategies [@Herault2011]. Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.  -->

# Material and Methods

## Study site

The study was carried out in a nearly 10-ha (300 m*330 m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996]. The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall of 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996]. The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

## Taxonomic composition and trait sampling

All trees $\geq$ 10 cm in diameter at breast height (dbh) have been identified to species level, tagged, mapped and monitored since 1989 following international standards [@Jairazbhoy2006], in an area progressively extended to a contiguous 10-ha plot in 2014.
The plot includes 6,672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot. We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density. 
<!-- Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively. -->

## Trait measurements

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$). Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an analytical balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.

## Environmental variables

We used airborne LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure. The LiDAR campaign was conducted in 2013, consistent with field measurements. Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher]. In addition, based on a 1-m resolution LiDAR model of canopy (Fig. \@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.

## Statistical Analyses

To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.

We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.

We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:

$$GWM_j=\sum_{i=1}^Sp_{i,j}*t_i$$
where $p_{i,j}$ was the relative abundance of species $i$ in subplot $j$, $t_i$ the mean trait value of species $i$ and $S$ the number of species in the guild considered.

We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild; [@bernard-verdier_community_2012]). The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances. We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as

$$SES = \frac{GWM_{obs}-\overline{GWM_{null}}}{\sigma_{null}}$$
where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.

We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions and a bidirectional stepwise selection of predictors. These analyses were performed independently for each of the five functional traits. All analyses were conducted using the R 3.3.0 software [@RCoreTeam2020].

# Results

## Trait variation and covariation

We found an overall increase of functional trait variation from intra-individual to inter-specific levels (Table \@ref(tab:traits)), even though LA showed almost as much variability at intra-individual and at intra-specific levels. Thus, trait values averaged at species level could grasp most of the functional variability across guilds and communities with varying taxonomic composition.

The first axis of the PCA performed on species trait values (Fig. \@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC. WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information. Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3 respectively, Fig. \@ref(fig:pca)B).

## Changes in functional composition with environment

At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig. \@ref(fig:GWMboxplot)). GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages. In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.

We detected a significant effect of slope, curvature and wetness on both canopy and understory guild assemblages, of southwesterness on canopy guild assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).
Wet and convex areas with more slope were associated to dense wood (WD) and thick (LT) leaves with high SLA in assemblages of both canopy and understory guilds. 
Convex areas showed opposed effect on LDMC, and wet areas opposed effect on LA, between assemblages of canopy and understory guilds.
Assemblages from canopy guilds under southwest orientation were associated to dense wood (WD) with small leaves (LA).
Assemblages from understory guilds under short canopy were associated to dense wood (WD) with thick (LT) and dense (LDMC) leaves with low SLA.

# Discussion

Few studies have investigated how the functional characteristics and composition of canopy and understorey guilds vary in tropical forests. 
We here show that functional trait values do not differ globally between canopy and understory tree species but that contrasting environments between understory and canopy filter species non-randomly, resulting in a markedly different functional composition of canopy and understory guilds within subplots.
We also showed that the guilds functional assemblages of canopy and understory  respond differently to environmental variation. Indeed, although the functional composition of understory and canopy guilds assemblages respond similarly to water and nutrient access, the functional composition of canopy guild communities strongly varies along the wood economics spectrum, in response to wind exposure, and the functional variation of understory assemblages mostly varies along the leaf economics spectrum, in response to  light availability. 
These results show that different environmental gradients simultaneously drives different functional responses among guilds in tropical rainforest.

<!-- We found that the measured functional traits of trees defined two orthogonal axes reflecting the leaf and wood economic spectra of trees at the scale of a 10ha rainforest plot. While functional trait values did not differ globally between canopy and understory tree species, we found neat difference between the assemblages of canopy and understory species within subplots. Therefore local environmental filtering selected species with diverging functional trait values across vertical guilds within communities. In addition the canopy and understory guild assemblages responded differently to environmental variation across communities. Soil resource, light availability and monsoon disturbances influenced the functional responses. Although understory and canopy assemblages responded similarly to water and nutrient access, functional variation of canopy assemblages was mostly along the wood economics spectrum, in response to aspect and wind exposure, while the functional variation of understory assemblages was mostly along the leaf economics spectrum in response to canopy height and decreasing light availability. These results show that environmental conditions drives different functional responses of vertical guilds in a tropical rainforest. -->

<!-- Assessing the variation of functional assemblages within tropical tree communities between understory and canopy species, we showed that while functional traits characteristics did not differ between species pools of canopy and understory, environmental filtering strongly favored different functional trait assemblages between canopy and understory guilds within communities. We further showed that the canopy and understory guilds assemblages responded differently across communities to environmental gradients related to soil resource, light availabilities and monsoon disturbances. While both understory and canopy assemblages responded similarly to water and nutrient access, canopy assemblages responded to monsoon winds with stronger support of the wood economics spectrum and understory assemblages responded to light access with a strong support of the leaf economics spectrum. These results illustrate how fine-scale environmental heterogeneity may shape the tree community composition and function. Moreover our study show guild-specific answer to biotic and abiotic environmental filtering between canopy and understory, highlighting the importance of forest vertical stratification in community assembly processes.  -->

## Decoupling of wood and leaf economics spectra at local scale

Functional traits PCA analysis supported two orthogonal main dimensions (Fig. \@ref(fig:pca)). 
<!-- Despite a reduced number of functional traits, our study support the existence of decoupled wood and leaf economics spectra at local scale in Indian rainforests. We found locally that species trait values varied along two orthogonal main dimensions (Fig. \@ref(fig:pca)).  -->
SLA and LDMC were opposed into a dimension representing the leaf economics spectrum [@diaz_global_2015; @wright_worldwide_2004], 
<!-- representing a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Baraloto2010; @Evans2001; @hodgson_is_2011]. -->
while, wood density was opposed to leaf area and thickness along a dimension reflecting the wood economics spectrum [@chave_towards_2009].
<!-- , which contrasted species more tolerant to hydric stress, with denser wood, to more competitive species with thicker and larger leaves and higher efficiency in water conductivity [@Santiago2004a].  -->
<!-- To our knowledge, our study is the first to support the two economics spectra in Indian rainforests. -->
@Baraloto2010 and @Fortunel2012a also found orthogonal dimensions reflecting wood and leaf economic spectra of Neotropical tree species at regional scale in Amazonia.
In response to these works, @Messier2017 suggested that these two orthogonal spectra reflect large scale difference in tree functional composition and would not be observed among species co-occurring at a local scale.
<!-- Here we show that the orthogonal spectra do exist across species at the scale of a 10ha plot. -->
Here we captured  the two orthogonal spectra across species at the scale of a 10 ha plot despite a limited number of traits compared to previous studies.
<!-- and (ii) we provide first evidence of these spectra in Indian rainforests. -->
Our results support that economics spectra are decoupled [@Baraloto2010] and not related to a single, whole-plant economics spectrum [@reich_world-wide_2014].

## Functional differences between canopy and understory guilds

Although the positions of canopy and understory species along the economics spectra did not differ globally (Fig. \@ref(fig:pca)B), we found significant trait differences between assemblages of canopy and understory species co-occuring in local 30 x 30m communities (Fig. \@ref(fig:GWMboxplot)). 
Therefore, species pools of canopy and understory species broadly overlap in functional spaces, and both spanned a broad range of ecological strategies (Fig. \@ref(fig:pca)A).
<!-- However, niche differentiation occurred in local communities across canopy and understory assemblages leading to contrasted functional compositions. -->
However, environmental niches filtered species in local communities leading to contrasted functional compositions between canopy and understory guild.
Understory assemblages had larger, thinner and lighter leaves, hence greater SLA, but denser wood, while canopy assemblages showed lighter wood but also smaller leaves. 
<!-- The differences did not support a mere differentiation of guild assemblages along economic spectra. -->
Light wood suggested greater hydraulic conduction of tall trees [@Santiago2018].
The conservative leaf strategy observed in canopy assemblage may respond to (i) tree height increasing leaf water stress due to gravity [@Koch2004], (ii) canopy light leading to a thicker mesophyll protecting tissues from excessive irradiance [@Chazdon1993], and (iii) a higher risk of herbivory and mechanical damage in canopy [@Sterck1992]. 
Drought constraint could be less important for understory trees, allowing greater investment into resource acquisition by leaves to solve the higher need for light in more shaded understory [@Chazdon1996].
However limited light may result in slower growth [@Herault2011] favoring denser wood. 
These results support the hypothesis that trees experience guild-specific ecological constraints within communities [@duque2002different], leading to contrasting functional compositions of canopy and understory assemblages. 
<!-- However the functional variation within guilds was still broad enough to entail overlap of the functional spaces at the scale of a 10ha plot, insuring greater stability in space and time to maintain the guilds functions [@Yachi1999]. -->

## Different responses to abiotic and biotic environment between canopy and understory assemblages

Our results show that the two guilds assemblages respond differently to environmental variation (Table \@ref(tab:LM)). 
Contrary to understory assemblages, canopy assemblages respond to southwesterness, which face dominant monsoon winds in our study site, along the wood economics spectrum with a greater WD and smaller LA for higher southwesterness.
The influence of aspect on tropical forest composition has been little documented.
@gunatilleke_specieshabitat_2006 found significant influence of aspect on forest composition in a similar climatic and topographical context in Sri Lanka and also interpreted it as a consequence of a higher exposure to monsoon winds.
In the Western Ghats of India, @Bunyan2015 found greater occurrence of forest fragments on southwest orientation in a forest-grassland mosaic, probably as a result from monsoon winds. 
The Uppangala plot is characterized by a large topographic heterogeneity, with steep slopes mostly organized along southwest-oriented crests (Fig. \@ref(fig:Uppangala)).
South-west oriented slopes are likely to be more exposed to monsoon winds, which could entail greater tree damages and forest disturbance [@soman_aspects_1990]. The consequences could be that (i) tree strategies associated with mechanical resistance and disturbance tolerance could be selected in areas more exposed to monsoon wind damages [@Poorter2010b], or (ii) turbulences on northeast oriented areas could yield more disturbance and death events favoring pioneer, lightwood species [@larjavaara_rethinking_2010].
In both cases, canopy species are more exposed to wind disturbances than understory trees and are thus expected to respond more to wind exposure.

Conversely, we found pervasive influence of canopy height on the functional composition of understory assemblages related to leaf economics, with more acquisitive strategies under tall canopies.
We indeed found a greater investment into photosynthetic activity, with higher SLA and lower LDMC and LT in understory assemblages under taller canopy. 
Taller canopies tend to bear greater vertical differentiation ensuring greater light availability and decreased investment in leaf lifespan in understory [@kitajima_tissue-level_2010].
Higher canopies filtered also smaller wood density for understory assemblages.
Lesser hydraulic constraints  in understory guilds could entail a greater role of competition among the trees, filtering lighter woods and relatively faster growth.

Nonetheless, curvature, wetness and slope similarly affected canopy and understory functional assemblages, except for LDMC. Locally convex wet, or flat areas tend to accumulate more soil, especially in hilly landscapes, and retain more water, increasing soil fertility [@ferry2010higher].
It thus create favorable conditions where species with faster resource acquisition (higher SLA, and lower LDMC and LT) and growth (lower WD and higher LA) strategies outcompete others.
By contrast, concave, drier or sloppy areas will be drier and less fertile favoring species assemblages with more conservative strategies and slower growth.

## Conclusion

Our results show that functional variations among species and assemblages preferentially vary along decoupled leaf and wood economic spectra in an Indian rainforest.
We found a significant environmental determinism in the functional composition of tree communities but also stressed that this determinism is guild-specific. Our results  thus contribute to the debate on the nature and role of economics spectra in the assembly of tropical forests [@Baraloto2010]. They also underline that the functional signature of community assembly should be decomposed to better characterize and understand the different strategies and responses across ecological guilds [@Raevel2018] and across forest vertical structure [@duque2002different]. We thus call for a better acknowledgment of differential guild dynamics within communities and their influence on the global forest functioning.

# Acknowledgements

We would like to thank Cedric Vega for extending Uppangala plot to 10 ha. We also want to acknowledge the data cleaning operated by Arianne Tanguy. And we are thankful to Claire Fortunel for her comments.

# Authors’ contributions

SS, MRM, FM and VR conceived the study, designed the methodology and analysed model outputs; SS, FM, VR, NA, NB and NB conducted the field sampling; NA, NB and NB identified species; MRM and CSJ provided the environmental data; SS, MRM, FM and VR led the writing of the manuscript. All the authors contributed to the drafts and gave their final approval for publication.

# Data availability

Functional traits data have been submitted to TRY initiative [@Kattge2019] under the name UppangalaTraits. Environmental variables and spatial positions of individuals were extracted from the Uppangala Station database or **lidar data**, for which access is modulated by **to precise**.

# References

<div id="refs"></div>

\newpage

## Table captions

__Table \@ref(tab:traits)__ Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).

__Table \@ref(tab:LM)__ Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$. See Table \@ref(tab:traits) for trait abbreviations.

## Figure captions

__Figure \@ref(fig:Uppangala)__ Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively.

__Figure \@ref(fig:pca)__ Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table \@ref(tab:traits) for trait abbreviations.

__Figure \@ref(fig:GWMboxplot)__ Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).

\newpage

<!-- table 1 -->

```{r traits, results="latex"}
t <- genPFT() %>% 
  mutate(id = row.names(.)) %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(id.vars = "id", variable.name = 'Abbreviation') %>% 
  mutate(Abbreviation = as.character(Abbreviation)) %>% 
  left_join(data.frame(
    Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
    Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
    Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
    Strategy = c('Leaf defense and investment',
                 'Leaf resource capture vs. investment',
                 'Leaf defense and investment',
                 'Leaf resource acquisition vs. defense',
                 'Stem transport, structure and defense'),
    stringsAsFactors = F
  ),  by = "Abbreviation") %>% 
  group_by(Trait, Abbreviation, Unit, Strategy) %>% 
  summarise_if(is.numeric, funs(min(.,na.rm = T), max(., na.rm = T))) %>% 
  mutate(range = paste(round(min,2), "-", round(max,2))) %>% 
  select(-min, -max) %>% 
  left_join(genPFT() %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'Abbreviation', 
                   value.name = 'Intra-individuals') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Tree, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'Abbreviation', value.name = 'Intra-species') %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Sp, Abbreviation) %>% 
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(Abbreviation) %>% 
              summarise_all(funs(mean(., na.rm = T))),  by = "Abbreviation") %>% 
  left_join(genPFT() %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_all(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>%
              mutate(id = row.names(.)) %>% 
              melt(id.vars = "id", 
                   variable.name = 'Abbreviation', value.name = 'Inter-species') %>% 
              select(-id) %>% 
              mutate(Abbreviation = as.character(Abbreviation)) %>% 
              group_by(Abbreviation) %>%
              summarise_all(funs(sd(., na.rm = T)/mean(., na.rm = T))),
            by = "Abbreviation") %>%
  mutate(`Intra-individuals` = ifelse(Abbreviation == "WD", NA, `Intra-individuals`))

if(knitr:::is_latex_output()){
  t %>% 
    ungroup() %>% 
    mutate(Trait = c('Leaf area', 'Leaf dry\nmatter content', 'Leaf\nthickness', 'Specific\nleaf area', 'Wood\ndensity'),
    Strategy = c('Leaf defense\nand investment',
                 'Leaf resource\ncapture vs. investment',
                 'Leaf defense\nand investment',
                 'Leaf resource\nacquisition vs. defense',
                 'Stem transport,\nstructure and defense')) %>% 
    mutate_all(linebreak) %>% 
    kable(caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).',
          digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
          col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "\\makecell[l]{Intra-\\\\individuals}", 
                        "\\makecell[l]{Intra-\\\\species}", 
                        "\\makecell[l]{Inter-\\\\species}")) %>%
    kable_styling(full_width = T)
} else {  
  t %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
        caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).') %>% 
  kable_styling(full_width = T)
}
```

\newpage

<!-- table 2 -->

```{r LM}
# GWM_ses %>% 
#   filter(!is.na(strata)) %>% 
#   melt(id.vars = c('strata', 'community')) %>% 
#   rename(trait = variable) %>% 
#   rename(SES = value) %>% 
#   mutate(community = as.character(community)) %>% 
#   left_join(R, by = "community") %>%
#   group_by(strata, trait) %>% 
#   do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#      direction = 'both', trace = 0)) %>% 
#   # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
#   #    direction = 'both', trace = 0) %>%
#   tidy(reg) %>%
#   group_by(strata, trait, term) %>%
#   # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
#                           # stars(p.value, ns = '   '))) %>% 
#   transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
#   dcast(strata + trait ~ term) %>% 
#   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
#   mutate_all(funs(replace(., is.na(.), ' '))) %>% 
#   left_join( # Inclusion of R2 but should be split
#     GWM_ses %>% 
#       filter(!is.na(strata)) %>% 
#       melt(id.vars = c('strata', 'community')) %>% 
#       rename(trait = variable) %>% 
#       rename(SES = value) %>% 
#       mutate(communtiy = as.character(community)) %>% 
#       left_join(R, by = "community") %>%
#       group_by(strata, trait) %>% 
#       do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#                     direction = 'both', trace = 0)) %>% 
#       glance(reg) %>% 
#       select(r.squared) %>% 
#       mutate(r.squared = round(r.squared, 2)) %>% 
#       rename("$R^2$" = r.squared), by = c("strata", "trait")
#   )%>% # End of R2
#   filter(trait != 'LES') %>% 
#   filter(trait != 'WES') %>%
#   # bold() %>% 
# Coefficients are unstable depending on random GWM so I entered the one from the last online version
data.frame(
  Guild = c(rep("C", 5), rep("U", 5)), Trait = rep(c("SLA", "WD", "LDMC", "LA", "LT"), 2),
  Slope = c(-9.5*10^-3, "", "", "", 1.3*10^-2, "", 1.3*10^-2, "", "", ""),
  Curvature = c(-2, 3.8, -3.2, -6, 3, -1.3, 6, 2.9, "", 6.6),
  Wetness = c(-0.16, "",  -0.28, "", "", -0.22, "", "", 0.32, 0.19),
  SW = c("", 0.23, "", -0.19, "", "", "", "", "", ""),
  Canopy = c("", "", "", 1.8*10^-2, "", 2.6*10^-2, -1.5*10^-2, -9*10^-3, "", -1.6*10^-2),
  R2 =c(0.06, 0.11, 0.04, 0.18, 0.05, 0.5, 0.32, 0.06, 0.04, 0.19)
) %>% 
  rename("Adjusted $R^2$" = R2) %>% 
  kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$ . See Table \\@ref(tab:traits) for trait abbreviations.', format = 'pandoc')
```

\newpage

<!-- figure 1 -->

```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
ggplot() +  
geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
    annotation_custom(grob = ggplotGrob(ggplot() +  
                                        geom_polygon(data = fortify(india),
                                                     aes(long, lat, group=group),
                                                     colour = "black", alpha = 0) +
                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
                                                   aes(long, lat), col = "red", size = 3) +
                                        coord_equal() +
                                        cowplot::theme_map() +
                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
                                              plot.background	= element_rect(fill = "white"))), 
                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
```

\newpage

<!-- figure 2 -->

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table  \\@ref(tab:traits) for trait abbreviations.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
# ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
#        file = "PCA.png",
#        path = './data/',
#        width = 200, height = 150, unit = 'mm',
#        dpi = 300)
```

\newpage

<!-- figure 3 -->

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>%
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  # select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  # facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
  # ggsave(file = "boxplot.png",
  #        path = './data/',
  #        width = 200, height = 150, unit = 'mm',
  #        dpi = 300)
```